\documentclass[12pt,preprint]{aastex631}
\usepackage{amsmath}

\begin{document}

\title{Mass-Dependent Radial Acceleration Relation}

\author{Anonymous}
\affiliation{Independent researcher}

\begin{abstract}
We present evidence that the Radial Acceleration Relation is better described by an interpolation function whose Weibull exponent varies with system mass: $p(M) = 2u/(1+3u)$, $u = (M/M_0)^{1/3}$. In a global fit with no per-galaxy fitting, this improves over the standard $p=0.5$ formula by 15.5\% in RMS ($\Delta\mathrm{BIC} = 38.4$). The coefficients $A=2$, $B=3$, $\alpha=1/3$, and $p_\mathrm{max}=2/3$ are all naturally determined by the data. We show that $p=0.5$ (McGaugh) follows from the flat rotation curve condition in 3D, and explore the connection to CDT spectral dimension flow.
\end{abstract}

\keywords{dark matter --- galaxies: kinematics and dynamics --- gravitation}

\section{Introduction}

The Radial Acceleration Relation (RAR) establishes a tight empirical correlation between observed gravitational acceleration $g_\mathrm{obs}$ and baryonic gravitational acceleration $g_\mathrm{bar}$ in disk galaxies \citep{McGaugh2016}. The standard interpolation function is
\begin{equation}
\mu(x) = 1 - \exp(-\sqrt{x}), \quad x = g_\mathrm{bar}/a_0
\end{equation}
where $a_0 \approx 1.2 \times 10^{-10}$ m/s$^2$. This corresponds to a Weibull distribution with shape parameter $p = 0.5$.

We find that allowing $p$ to vary with system mass $M$ significantly improves the fit:
\begin{equation}
p(M) = \frac{2u}{1+3u}, \quad u = \left(\frac{M}{M_0}\right)^{1/3}
\label{eq:pM}
\end{equation}

\section{Data and Methods}

We use 175 SPARC galaxies \citep{Lelli2016} and 17 galaxy clusters from \citet{Vikhlinin2006} and X-COP \citep{Ettori2019}. The global fit uses a single $Y_\mathrm{disk}$ for all galaxies (no per-galaxy fitting), addressing the \citet{Li2021} bias concern.

\section{Results}

\subsection{Global Fit}

\begin{deluxetable}{lccc}
\tablecaption{Model Comparison\label{tab:models}}
\tablehead{\colhead{Model} & \colhead{Params} & \colhead{RMS} & \colhead{vs McGaugh}}
\startdata
McGaugh ($p=0.5$) & 1 & 0.197 & baseline \\
Constant $p$ & 2 & 0.195 & $+0.9\%$ \\
Variable $G(M)$ [control] & 3 & 0.188 & $+4.6\%$ \\
\textbf{p(M) model} & \textbf{3} & \textbf{0.166} & $\mathbf{+15.5\%}$ \\
\enddata
\end{deluxetable}

The $\Delta\mathrm{BIC} = 38.4$ between McGaugh and $p(M)$ with $\alpha=1/3$ constitutes very strong evidence for the mass-dependent model.

\subsection{Parameter Precision}

Fixing $\alpha = 1/3$ degrades the fit by only 0.02\% compared to the free fit ($\alpha = 0.312$). The asymptotic value $p_\mathrm{max} = A/B = 2/3$ matches the free fit to 0.00\%. The coefficients $A = 2.015$, $B = 3.012$ are determined to 0.5\% precision.

\subsection{Theoretical Derivation}

The condition for flat rotation curves in $d$ spatial dimensions gives
\begin{equation}
p_\mathrm{flat} = \frac{d-2}{d-1}
\end{equation}
For $d=3$: $p_\mathrm{flat} = 1/2$, recovering the McGaugh value from first principles.

\subsection{CDT Connection}

We implement 2+1D and 3+1D Causal Dynamical Triangulations in C with full Pachner moves. The 2+1D simulation (L=30, 250k tetrahedra, 5-seed average) yields $\gamma = 0.87 \pm 0.1$ for the spectral dimension flow, consistent with $\gamma \approx 1$ (Ambjorn formula). The mathematical identity
\begin{equation}
\frac{2u}{1+3u} \equiv \frac{2}{3} \cdot \frac{v}{1+v}, \quad v = 3u
\end{equation}
shows that CDT $\gamma = 1$ is compatible with galaxy $\alpha = 1/3$ through the geometric mapping $\sigma \sim r \sim M^{1/3}$.

The 3+1D CDT simulation (L=8, 147k four-simplices, all three Pachner move types (2,4)/(4,2)/(3,3)) yields $d_\mathrm{spec}(\sigma=10) = 3.5$--$3.7 < 4$, consistent with UV dimensional reduction. The UV spectral dimension is stable across Regge action parameters ($k_4 = 0.5$--$3$).

\subsection{Galaxy Morphology}

Disk-dominated galaxies (B/T $< 0.1$) give $\alpha = 0.341 \approx 1/3$ with $+7.3\%$ improvement, while bulge-dominated galaxies give $\alpha = 0.264$ with only $+0.5\%$ improvement (McGaugh already sufficient).

\section{Discussion}

The mass-dependent RAR exponent is an observational finding supported by $\Delta\mathrm{BIC} = 38.4$. The formula $p(M) = 2u/(1+3u)$ has all coefficients naturally determined by the data. The CDT connection provides a mathematical framework but not additional predictions, as $\gamma$ is absorbed into $M_0$.

Leave-one-out cross-validation within SPARC gives $+5.4\%$ improvement, confirming the model is not overfitting. The 3+1D CDT simulation yields $d_\mathrm{spec}(\sigma=10) = 3.5$--$3.7$, showing UV dimensional reduction below $d=4$.

Key limitations: (1) SPARC is the only galaxy dataset tested; (2) the Li+ gap with baryonic masses (6.3\%) is smaller than with dynamical masses (11\%); (3) $M_0$ and $a_0$ have no fundamental derivation.

\section{Conclusion}

The RAR interpolation exponent depends on system mass. The specific form is determined to sub-percent precision: $\alpha = 1/3$, $p_\mathrm{max} = 2/3$, $A = 2$, $B = 3$. Independent verification with non-SPARC data is the critical next step.

\begin{acknowledgments}
Code and data: \url{https://github.com/uuki555-cyber/sdhg-gravity}
\end{acknowledgments}

\bibliography{references}

\end{document}
